---
title: "Motor Trend Analysis"
author: "Cary Correia"
date: "November 7, 2014"
output: pdf_document
---

Context

You work for Motor Trend, a magazine about the automobile industry. Looking at a data set of a collection of cars, they are interested in exploring the relationship between a set of variables and miles per gallon (MPG) (outcome). They are particularly interested in the following two questions:

“Is an automatic or manual transmission better for MPG”
"Quantify the MPG difference between automatic and manual transmissions"

Executive Summary


```{r}
summary(cars)
```

You can also embed plots, for example:

```{r, echo=FALSE}
plot(cars)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

Appendix:

Check off list for grading:
- did the student interpret the coefficients correctly?

- did the student fit multiple models and detail their strategy for model selection ?
- Did the student answer the questions of interest or detail why the question(s) is (are) not answerable?
- Did the student do a residual plot and some diagnostics?
- Did the student quantify the uncertainty in their conclusions and/or perform an inference correctly?
-Was the report brief (about 2 pages long) for the main body of the report and no longer than 5 with supporting appendix of figures?
-Did the report include an executive summary?]
-Was the report done in Rmd (knitr)?

Motor Trend Car Road Tests

Description

The data was extracted from the 1974 Motor Trend US magazine, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973–74 models).

Usage

mtcars
Format

A data frame with 32 observations on 11 variables.

[, 1]   mpg	Miles/(US) gallon
[, 2]	cyl	Number of cylinders
[, 3]	disp	Displacement (cu.in.)
[, 4]	hp	Gross horsepower
[, 5]	drat	Rear axle ratio
[, 6]	wt	Weight (lb/1000)
[, 7]	qsec	1/4 mile time
[, 8]	vs	V/S
[, 9]	am	Transmission (0 = automatic, 1 = manual)
[,10]	gear	Number of forward gears
[,11]	carb	Number of carburetors
